### 2017-03-15 main.tex
Textual updates for the plethora of new birds.

With inputs well outside the space on which a neural network has been
trained, its outputs will be essentially random. In order to reduce
the false positive rate it is necessary to provide negative training
examples that include silence, cage noise, wing flapping, non-song
vocalisations, and perhaps songs from other birds.  Although it will
depend on the makeup of the non-song data, we have found that training
with as low as a 1:1 ratio of non-song to song---or roughly 10-20
minutes of non-song audio---yields excellent results on most birds.
These data are also included in our accuracy tests.

@@

We used an FFT of size 256; a Hamming window; and chose a target
spectrogram frame interval of $t_\textrm{fft}=1.5$ milliseconds,
resulting in a true frame interval of $t_\textrm{fft}=\lfloor 1.5\cdot
44.1\rceil /44.1\approx 1.4966$ ms.  We set the network's input space
to 30 ms long, and to span frequencies from 1--8 kHz, which contains
the fundamentals and several overtones of most zebra finch
vocalisations.

@@

We measure accuracy and characterise timing for 6 randomly chosen
points in one bird's song (Bird A), and present accuracy results for
three more birds (B, C, D) (timing does not appreciably vary among
birds). The results presented here are representative of real-world
performance in feedback experiments on many birds in our lab.

% A:
% B: LNY46 0.28s @ 140ms
% C: LNY42 0.38s
% D: LNY4RB 0.25s
% E: LR28 0.14s

\begin{table}
 \begin{tabular}{c|cccccc}
    Bird & Training songs & Test songs & Training non-song & Test non-song & Target time  & Time window length\\ %\hline
   A & 1000 & 1000 & 1000 & 3636 & 150:50:400 & 50 \\
   B & 1000 & 1000 & 1000 & 1000 & 195 & 50 \\
    C & 1000 & 1000 & 1000 & 1000 & 380 & 50 \\
    D & 1000 & 1000 &  1000 & 1000 & 250 & 50 \\
    E & 1000 & 1000 & 1000 & 1000 & 140 & 80
  \end{tabular}
  \caption{Sizes of our training and test sets, times of interest, and detection time window length (milliseconds).}
  \label{table:datasizes}
\end{table}

Data consisted of extracted, aligned songs, and additional song-length
samples of non-song (cage noise, calls, other birds' song). These were
divided into training and test sets. For all birds, we used 1000 songs
and 1000 non-song samples for training. For Bird A we used the
remaining 1818 songs and 3636 nonsongs for testing. For each of the
other birds, we tested on another 1000 songs and 1000
nonsongs. Table~\ref{table:datasizes} lists the dataset sizes, along
with the target detection points and the length of the time history
window used for detection, which was the only parameter that we
adjusted.

@@

\fig{fig:song} shows the song for which we present accuracy and timing
results.  Six example target trigger points are shown, at 150, 200,
250, 300, 350, and 400 milliseconds after the beginning of the aligned
samples, as indicated by the red lines, and referred to henceforth as
$t^*_1\ldots t^*_6$ respectively.  The rectangles show the recognition
window (30 ms, 1--8 kHz) for each instant in the song.

@@

\begin{figure}
  \includegraphics[width=\textwidth]{Fig2}
  % 9x10 inches
  % Fonts: Helvetica, scaled 130%
  \caption{Each plot shows one network output unit's responses to all
    2818 presentations of the song shown in \fig{fig:song}.  For
    simplicity, we show only the syllables $t^*_1$, $t^*_4$, and
    $t^*_6$, and we do not show the non-response to presentation of non-song. The horizontal axis is time relative to the beginning of
    the aligned song, and the vertical axis is an index for the 2818
    individual song presentations. The grey shading shows the audio
    amplitude of song Y at time T. Detection events on training songs
    are shown in cyan, with detections of unseen test songs in red.  To provide an intuition of intra-song variability, songs have been
    stably sorted by the time of detection events; thus, each of the
    three detection graphs shows the songs in a different order.}
  \label{fig:detection_raster}
\end{figure}

@@

Accuracies for our six test syllables for bird A are shown in
\fig{fig:accuracies} (accuracies for the synthetic $\delta$-function
songs were always 100\%).  For each test syllable we trained 100
different networks, which differ from each other in the random
initialisation of network weights and on which random subset of our
songs was used for training.  Each network's accuracy is shown as a
single point in each chart in \fig{fig:accuracies}, and the means are
shown in Table~\ref{table:accuracies}.  Evidently some syllables are
easier to identify reliably than others (for example, $t^*_3$ and
$t^*_6$ are occasionally confused due to their similar appearance, as
could be guessed from \fig{fig:song}); thus, if the experimental
design permits, it can be worthwhile to choose an easily identifiable
syllable.  Furthermore, as can be seen from \fig{fig:accuracies}, for
difficult syllables the training process occasionally yields a network
that performs unusually poorly.  It is easy to spot these networks by
their performance either as measured as part of the training process
or on the test audio file, perhaps taking the best of a few training
runs.

@@

  \caption{Accuracy variability over 100 different training runs for
    each of the six test syllables.  Each dot shows the test-set
    accuracy for an independently trained detector.  Because the
    horizontal positions have been randomised slightly so as not to
    occlude same-valued measurements, test syllable is also indicated
    by colour.  The means are given in Table~\ref{table:accuracies}.}
  \label{fig:accuracies}
\end{figure}

@@

Detector latency and jitter depend on the FFT frame rate.  Our
1.5-ms--frame default is a compromise: shorter frames increase the
precision of timing, but also increase computational requirements both
during training and at runtime.  \fig{fig:TimingVsFrame} shows how
these numbers vary over a useful range of frame rates on our ideal
detector, the Swift detector with serial output, and for the LabVIEW
detector, for both the $\delta$-syllable and for the bird song at
$t^*_4$.

\begin{figure}
  \includegraphics[width=\textwidth]{Fig4}
  \caption{Timing varies as the FFT frame interval changes.  Here we
    show results for the ideal detector and the LabVIEW and
    Swift+serial implementations, for the constructed
    $\delta$-syllable and for trigger $t^*_4$ of the bird song.  The
    lines show latency; error bars are standard deviation (jitter).  Points have
    been shifted horizontally slightly for clarity; original positions are [0.5 1
      1.5 2 4] ms.}
  \label{fig:TimingVsFrame}
\end{figure}

@@

Syllable choice impacts detector performance, but despite the variety
of syllables chosen here, performance was fairly stable across
syllables.  \fig{fig:TimingVsSyllable} shows measured timing data for
the Swift+serial detector compared to the ideal, with
$t_{\textrm{fft}}=1.5$ ms.  Given the variety of the test syllables,
the variability is surprisingly low, and is summarised here with 95\%
confidence intervals (with n=6):
\vspace{8pt}\par\noindent
\begin{tabular}{r|cc}
  Detector & Latency & Jitter \\ 
  \hline   Ideal & $-0.8\pm 0.3$ ms & $2.1\pm 0.1$ ms \\
  Swift+serial & $0.0\pm 0.6$ ms & $2.0\pm 0.1$ ms
\end{tabular}
\vspace{8pt}\par\noindent
The negative latency is due to the way in which the network responds
to the song through time: as the recognition region looks increasingly
similar to the trained match, the network's evaluation of similarity
rises, and will generally cross the triggering threshold before it
reaches its maximum value.  A heuristic as simple as triggering at an
apparent turning point after crossing the threshold might improve
timing consistency at the expense of latency, but we did not test this.

\begin{figure}
  \begin{center}
    \includegraphics[width=8cm]{Fig5}
  \end{center}
  \caption{Timing data for our 6 test syllables, for the ideal and the
    Swift+serial detectors, with an FFT frame rate of 1.5 ms.  Point
    centres show latency; error bars show jitter.}
  \label{fig:TimingVsSyllable}
\end{figure}

@@

We compared latency and jitter across our different detector
implementations for the $\delta$-syllable and $t^*_4$, again with
$t_\mathrm{fft}=1.5$ ms.  Results are as follows:
\vspace{8pt}
\begin{tabular}{l|c|c|c| c}
  & \multicolumn{2}{c}{$\delta$-syllable} & \multicolumn{2}{c}{Bird: $t^*_4$} \\
  Detector & Latency (ms) & Jitter (ms) & Latency (ms) & Jitter (ms) \\
  \hline
  Ideal & 0.66 & 0.38 & -1.0 & 2.0 \\
  LabVIEW & 1.5 & 0.69 & -0.18 & 2.4 \\
  Swift+serial & 0.88 & 1.6 & -0.075 & 2.0 \\
  Swift+audio & 5.9 & 2.6 & 1.2 & 4.9 \\
  MATLAB+serial & 8.1 & 2.3 & 3.8 & 2.0 \\
  MATLAB+audio & 13 & 5.9 & 8.4 & 5.9
\end{tabular}
\vspace{8pt}\par\noindent
These data are also plotted in \fig{fig:TimingVsDetector}, and
\fig{fig:timing} gives a more detailed view of what the timing curves
look like for the five implementations of our detector on $t^*_4$.

\begin{figure}
  \begin{center}
    \includegraphics[width=8cm]{Fig6}
  \end{center}
  \caption{The different detectors for the constructed
    $\delta$-syllable and for the bird song at $t^*_4$.  Point centres
    show latency; error bars show jitter.}
  \label{fig:TimingVsDetector}
\end{figure}

\begin{figure}
  \begin{center}
    \includegraphics[width=8cm]{Fig7}
  \end{center}
  \caption{Raw timing curves for all detectors measured during
    detection of $t^*_4$ using 1.5-ms frames.  We extract the trigger events from each curve, from which we obtain the mean---latency---and
    standard deviation---jitter.}
  \label{fig:timing}
\end{figure}

@@

The song presented here was recorded with a fixed microphone mounted
inside the cage.  We found that higher accuracy is achieved when a
microphone is instead mounted on the bird's head, which maintains
volume and reduces changes in timbre as the bird moves.
% We have head mounted microphone data?

>>>>

With inputs well outside the space on which a neural network has been
trained, its outputs will be essentially random. In order to reduce
the false positive rate it is necessary to provide negative training
examples that include silence, cage noise, wing flapping, non-song
vocalisations, and perhaps songs from other birds.  Although it will
depend on the makeup of the non-song data, we have found that training
with as low as a 1:1 ratio of non-song to song---or roughly 10
minutes of non-song audio---yields excellent results on most birds.
These data are also included in our accuracy tests.

@@

We used an FFT of size 256; a Hamming window; and chose a target
spectrogram frame interval of $t_\textrm{fft}=1.5$ milliseconds,
resulting in a true frame interval of $t_\textrm{fft}=\lfloor 1.5\cdot
44.1\rceil /44.1\approx 1.4966$ ms.  We set the network's input space
to 50 ms long, and to span frequencies from 1--8 kHz, which contains
the fundamentals and several overtones of most zebra finch
vocalisations.

@@

We measure accuracy and characterise timing for 6 randomly chosen
points in one bird's song (lny64), and present accuracy results for
seven more birds (the detector's timing does not appreciably vary among
birds). The results presented here are representative of real-world
performance in feedback experiments on many birds in our lab.

\begin{table}
 \begin{tabular}{c|ccc}
    Bird & Test songs & Test non-songs & Target time (ms) \\ \hline
    lny64 & 1818 & 10545 & 150:50:400 \\
    lny46 & 1000 & 1000 & 195 \\
    lny42 & 1245 & 1223 & 380 \\
    lny4rb & 1000 & 1000 & 250 \\
    lr12 & 2000 & 2008 & 180 \\
    lr13 & 2000 & 2001 & 300 \\
    lr28 & 3000 & 3014 & 140 \\
    lr77 & 2000 & 2008 & 320
  \end{tabular}
  \caption{For each bird, detectors were trained for the specified
    timepoints, using 1000 songs and 1000 non-song samples of the same
    length. The rest of the audio samples---shown here---were used to
    test the learned detectors.}
  \label{table:datasizes}
\end{table}

Data consisted of extracted, aligned songs, and additional song-length
samples of non-song (cage noise, calls, other birds' song). These were
divided into training and test sets. For all birds, we used 1000 songs
and 1000 non-song samples for training. Table~\ref{table:datasizes}
lists the test dataset sizes and target detection points. For lny64,
we arbitrarily chose six linearly spaced target times. For each of the
other birds, we eyeballed one promising-looking detection point.

@@

\fig{fig:song} shows a typical song. The image is produced by
averaging the aligned spectrograms, in this case for the bird lny64.
Six target trigger points are shown for this song, at 150, 200, 250,
300, 350, and 400 milliseconds after the beginning of the aligned
samples, as indicated by the red lines, and referred to henceforth as
$t^*_1\ldots t^*_6$ respectively.  For clarity, the rectangles show a
shorter recognition window than our standard: here 30 ms, but still
spanning 1--8 kHz, for each target time in the song.\marginpar{Is this
  OK?  Should I update to a 50-ms window? They'll overlap!}

@@

\begin{figure}
  \includegraphics[width=\textwidth]{Fig2}
  % 9x10 inches
  % Fonts: Helvetica, scaled 130%
  \caption{Each plot shows one network output unit's responses to all
    2818 presentations of lny64's song shown in \fig{fig:song}.  For
    simplicity, we show only the syllables $t^*_1$, $t^*_4$, and
    $t^*_6$, and we do not show the non-response to presentation of non-song. The horizontal axis is time relative to the beginning of
    the aligned song, and the vertical axis is an index for the 2818
    individual song presentations. The grey shading shows the audio
    amplitude of song Y at time T. Detection events on training songs
    are shown in cyan, with detections of unseen test songs in red.  To provide an intuition of intra-song variability, songs have been
    stably sorted by the time of detection events; thus, each of the
    three detection graphs shows the songs in a different order.}
  \label{fig:detection_raster}
\end{figure}

@@

Accuracies for our six test syllables for bird lny64 are shown
in\marginpar{Update if I include all birds in this figure!}
\fig{fig:accuracies} (accuracies for the synthetic $\delta$-function
songs were always 100\%).  For each test syllable we trained 100
different networks, which differ from each other both in the random
initialisation of network weights and in which random subset of the
song-and-nonsong corpus was used for training.  Each network's
accuracy is shown as a single point in each chart in
\fig{fig:accuracies}, and the means are shown in
Table~\ref{table:accuracies}.  Evidently some syllables are easier to
identify reliably than others (for example, $t^*_3$ and $t^*_6$ are
occasionally confused due to their similar appearance, as could be
guessed from \fig{fig:song}); thus, if the experimental design
permits, it can be worthwhile to choose an easily identifiable
syllable.  Furthermore, as can be seen from \fig{fig:accuracies}, for
difficult syllables the training process occasionally yields a network
that performs unusually poorly.  It is easy to spot these networks by
their performance either as measured as part of the training process
or on the test audio file, perhaps taking the best of a few training
runs.

@@

  \caption{Accuracy variability over 100 different training runs for
    each of the six test syllables for lny64.  Each dot shows the test-set
    accuracy for an independently trained detector.  Because the
    horizontal positions have been randomised slightly so as not to
    occlude same-valued measurements, test syllable is also indicated
    by colour.  The means for all birds are given in Table~\ref{table:accuracies}.}
  \label{fig:accuracies}
\end{figure}

@@

Detector latency and jitter depend on the FFT frame rate.  Our
1.5-ms--frame default is a compromise: shorter frames increase the
precision of timing, but also increase computational requirements both
during training and at runtime.  \fig{fig:TimingVsFrame} shows how
these numbers vary over a useful range of frame rates on our ideal
detector, the Swift detector with serial output, and for the LabVIEW
detector, for both the $\delta$-syllable and for lny64's song at
$t^*_4$.

\begin{figure}
  \includegraphics[width=\textwidth]{Fig4}
  \caption{Timing varies as the FFT frame interval changes.  Here we
    show results for the ideal detector and the LabVIEW and
    Swift+serial implementations, for the constructed
    $\delta$-syllable and for trigger $t^*_4$ of lny64's song.  The
    lines show latency; error bars are standard deviation (jitter).  Points have
    been shifted horizontally slightly for clarity; original positions are [0.5 1
      1.5 2 4] ms.}
  \label{fig:TimingVsFrame}
\end{figure}

@@

Syllable choice impacts detector performance, but despite the variety
of syllables shown here, performance was stable across syllables.
\fig{fig:TimingVsSyllable} shows measured timing data for lny64's test
syllables for the Swift+serial detector compared to the ideal, with
$t_{\textrm{fft}}=1.5$ ms.  Given the variety of the test syllables,
the variability is surprisingly low, and is summarised here with 95\%
confidence intervals (with n=6):
\vspace{8pt}\par\noindent
\begin{tabular}{r|cc}
  Detector & Latency & Jitter \\ 
  \hline   Ideal & $-0.8\pm 0.3$ ms & $2.1\pm 0.1$ ms \\
  Swift+serial & $0.0\pm 0.6$ ms & $2.0\pm 0.1$ ms
\end{tabular}
\vspace{8pt}\par\noindent
The negative latency is due to the way in which the network responds
to the song through time: as the recognition region looks increasingly
similar to the trained match, the network's evaluation of similarity
rises, and will generally cross the triggering threshold before it
reaches its maximum value.  A heuristic as simple as triggering at an
apparent turning point after crossing the threshold might improve
timing consistency at the expense of latency, but we did not test this.

\begin{figure}
  \begin{center}
    \includegraphics[width=8cm]{Fig5}
  \end{center}
  \caption{Timing data for lny64's 6 test syllables, for the ideal and the
    Swift+serial detectors, with an FFT frame rate of 1.5 ms.  Point
    centres show latency; error bars show jitter.}
  \label{fig:TimingVsSyllable}
\end{figure}

@@

We compared latency and jitter across our different detector
implementations for the $\delta$-syllable and lny64's $t^*_4$, again with
$t_\mathrm{fft}=1.5$ ms.  Results are as follows:
\vspace{8pt}
\begin{tabular}{l|c|c|c| c}
  & \multicolumn{2}{c}{$\delta$-syllable} & \multicolumn{2}{c}{Bird lny64: $t^*_4$} \\
  Detector & Latency (ms) & Jitter (ms) & Latency (ms) & Jitter (ms) \\
  \hline
  Ideal & 0.66 & 0.38 & -1.0 & 2.0 \\
  LabVIEW & 1.5 & 0.69 & -0.18 & 2.4 \\
  Swift+serial & 0.88 & 1.6 & -0.075 & 2.0 \\
  Swift+audio & 5.9 & 2.6 & 1.2 & 4.9 \\
  MATLAB+serial & 8.1 & 2.3 & 3.8 & 2.0 \\
  MATLAB+audio & 13 & 5.9 & 8.4 & 5.9
\end{tabular}
\vspace{8pt}\par\noindent
These data are also plotted in \fig{fig:TimingVsDetector}, and
\fig{fig:timing} gives a more detailed view of what the timing curves
look like for the five implementations of our detector on $t^*_4$.

\begin{figure}
  \begin{center}
    \includegraphics[width=8cm]{Fig6}
  \end{center}
  \caption{The different detectors for the constructed
    $\delta$-syllable and for lny64's song at $t^*_4$.  Point centres
    show latency; error bars show jitter.}
  \label{fig:TimingVsDetector}
\end{figure}

\begin{figure}
  \begin{center}
    \includegraphics[width=8cm]{Fig7}
  \end{center}
  \caption{Raw timing curves for all detectors measured during
    detection of lny64's $t^*_4$ using 1.5-ms frames.  We extract the trigger events from each curve, from which we obtain the mean---latency---and
    standard deviation---jitter.}
  \label{fig:timing}
\end{figure}

@@

The songs presented here were recorded with fixed microphones mounted
inside the cages.  We found that higher accuracy is achieved when a
microphone is instead mounted on the bird's head, which maintains
volume and reduces changes in timbre as the bird moves.
% We have head mounted microphone data?
